---
title: "Tutorial analyses MT dataset"
author: "Pascal J. Kieslich & Michael Schulte-Mecklenbeck"
output:
  ioslides_presentation
---

# Preparations

## Load libraries
```{r,message=FALSE}
library(mousetrap)
library(ggplot2)
library(dplyr)
```

If you have not installed these packages yet, you have to run `install.packages()` first, so e.g., `install.packages('mousetrap')` to install the mousetrap R package.

- `ggplot` will give us nice plotting options.
- `dplyr` is the ultimate swissaRmy-knife for data handling.

```{r, eval=F, echo=F}
theme_set(theme_classic()+ 
  theme(
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    axis.text = element_text(colour = "black"),
    panel.border = element_rect(colour = "black", fill=NA)
  ))
```

# Import + Processing

## Import
```{r}
mt_data <- mt_import_mousetrap(KH2017_raw)
```

We first load the KH2017_raw dataset, that comes with the `mousetrap` package.

## Spatial transformations
```{r}
# Remap trajectories
mt_data <- mt_remap_symmetric(mt_data)

# Align trajectories to common start position
mt_data <- mt_align_start(mt_data, start=c(0,0))
```

## Resampling
```{r}
# Time-normalize trajectories
mt_data <- mt_time_normalize(mt_data) 
```

## Filtering
```{r}
# Only keep trials with correct answers
mt_data <- mt_subset(mt_data, correct==1)
```

# Analysis

## Aggregate trajectories
```{r message=F, out.width='80%'}
mt_plot_aggregate(mt_data, use="tn_trajectories",
  x="xpos", y="ypos", 
  color="Condition", subject_id="subject_nr")+
  scale_color_manual(values=c("darkorange","steelblue"))
```

## Calculate measures
```{r}
# Calculate velocity and acceleration
mt_data <- mt_derivatives(mt_data)

# Calculate trial-level indices
mt_data <- mt_measures(mt_data)

# Calculate sample entropy
mt_data <- mt_sample_entropy(mt_data, use="tn_trajectories")
```


## Curvature - Aggregate analyses
```{r message=FALSE}
# Aggregate MAD values per participant and condition
agg_mad <- mt_aggregate_per_subject(mt_data, 
  use_variables="MAD", 
  use2_variables="Condition",
  subject_id="subject_nr")
```

## Curvature - Aggregate analyses
```{r}
# Compare aggregated MAD values
t.test(MAD ~ Condition, data=agg_mad, paired=TRUE)
```

## Curvature - Aggregate analyses
```{r}
# Calculate descriptives
agg_mad %>% 
  group_by(Condition) %>% 
  summarise_at("MAD",.funs=c("mean","sd"))
```

## Heatmap of individual trajectories
```{r heatmap_plot, echo = TRUE, eval = FALSE}
mt_heatmap(mt_data,
           xres=1000,
           colors=c("white","black"))
```

## Heatmap of individual trajectories
```{r heatmap_plot, echo = FALSE, eval = TRUE, out.width='70%'}
```

## Smoothed heatmap
```{r smoothed_heatmap_plot, echo = TRUE, eval = FALSE}
mt_heatmap(mt_data,
           xres=1000,
           colors=c("white","black"),
           smooth_radius=20,
           n_shades=10,
           mean_image=0.2)
```

## Smoothed heatmap
```{r smoothed_heatmap_plot, echo = FALSE, eval = TRUE, out.width='70%'}
```

## Diff heatmap between conditions
```{r diff_heatmap_plot, echo = TRUE, eval = FALSE}
mt_diffmap(mt_data,
           xres=1000,
           condition=mt_data$data$Condition=="Typical",
           colors=c("steelblue","white","darkorange"),
           smooth_radius=20,
           n_shades=10)
```

## Diff heatmap between conditions
```{r diff_heatmap_plot, echo = FALSE, eval = TRUE, out.width='70%'}
```

## Temporal analyses

### Average x positions
```{r, message=F, out.width='70%'}
# Plot aggregate time-normalized x-positions (Fig. 6)
mt_plot_aggregate(mt_data, use="tn_trajectories",
                  x="steps", y="xpos", color="Condition",
                  subject_id="subject_nr", points=TRUE)+
  scale_color_manual(values=c("darkorange","steelblue"))

# Aggregate time-normalized trajectories per condition
# separately per participant
av_tn_trajectories <- mt_aggregate_per_subject(mt_data,
                                               use="tn_trajectories", use2_variables="Condition",
                                               subject_id="subject_nr")

# Paired t-tests on coordinates
xpos_t_tests <- 
  with(av_tn_trajectories,
       sapply(unique(steps),function(i){
         t.test(xpos[Condition=="Atypical" & steps==i],
                xpos[Condition=="Typical" & steps==i],
                paired = TRUE)$p.value})
  )

# Retrieve all significant t-tests
which(xpos_t_tests<.05)
```

## Riverbed plot
```{r riverbed, echo = TRUE, eval = FALSE, out.width='70%'}
mt_plot_riverbed(mt_data, use="tn_trajectories",
                 y="xpos", facet_col="Condition")
```

## Riverbed plot
```{r riverbed, echo = FALSE, eval = TRUE, out.width='90%'}
```

# Summary

